Evolution Reinforces Cooperation with the Emergence of Self-Recognition Mechanisms: an Empirical Study of Strategies in the Moran process for the Iterated Prisoner's Dilemma

### 2018-04-18 main.tex

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline
{Evolution reinforces cooperation with the emergence of self-recognition
 mechanisms: an empirical study of strategies in the Moran process for the
 iterated prisoner's dilemma }
% Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Vincent Knight\textsuperscript{1*\Yinyang},
Marc Harper\textsuperscript{2\Yinyang},
Nikoleta E. Glynatsi\textsuperscript{2\ddag},
Owen Campbell\textsuperscript{3\ddag},
\\
\bigskip
\textbf{1} Cardiff University, School of Mathematics, Cardiff, UK
\\
\textbf{2} Google Inc., Mountain View, CA, USA
\\
\textbf{3} Independent Researcher, Chester, UK
\\
\bigskip

@@

\begin{enumerate}
 \item Zero determinant strategies are not effective invaders or defenders for $N > 2$
 \item Complex strategies can be effective, and in fact can naturally evolve
     through evolutionary processes to outperform designed strategies.
 \item The strongest resistors specifically evolve or possess a handshake mechanism.
 \item Strong invaders are generally cooperative strategies that do not defect
 first but retaliate to varying degrees of intensity against strategies that defect.
 \item Strategies evolved to maximize their total payoff can be strong invaders
 and achieve mutual cooperation with many other strategies.
\end{enumerate}

@@

All of the training code is archived at \cite{marc_harper_2017_824264}. This
software is (similarly to the Axelrod library) available on
github\footnote{\url{https://github.com/Axelrod-Python/axelrod-dojo}}
with documentation to
train new strategies easily. Training
typically takes less than 100 generations and can be completed within several
hours on commodity hardware.

There are three further strategies trained specifically for this study; Trained
FSM 1, 2, and 3 (TF1 - TF3). These are based on finite state machines of 16, 16,
and 8 states respectively. These are shown in Fig~\ref{fig:tf1},~\ref{fig:tf2}
and~\ref{fig:tf3}), using the notation common in the literature where
\(A_1/A_2\) is the action of the opponent \(A_1\) and the response of the player
\(A_2\) as well as arrows corresponding to changes of state.

@@

As will be demonstrated the results for
\(N=2\) differ from those of larger $N$. Hence these results do not concur with
the literature which suggests that zero determinant strategies should be
effective for larger population sizes, but these analyses consider stationary
behaviour, while this work runs for a fixed number of rounds. \cite{stewart2013extortion}
The stationarity assumption allows for a deterministic payoff matrix
leading to the conclusions about zero determinant strategies in the space
of memory-one strategies that do not generalize to this context.

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline
{Evolution reinforces cooperation with the emergence of self-recognition
 mechanisms: an empirical study of strategies in the Moran process for the
 iterated prisoner's dilemma }
% Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Vincent Knight\textsuperscript{1*\Yinyang},
Marc Harper\textsuperscript{2\Yinyang},
Nikoleta E. Glynatsi\textsuperscript{1\ddag},
Owen Campbell\textsuperscript{3\ddag},
\\
\bigskip
\textbf{1} Cardiff University, School of Mathematics, Cardiff, UK
\\
\textbf{2} Google Inc., Mountain View, CA, USA
\\
\textbf{3} Independent Researcher, Chester, UK
\\
\bigskip

@@

\begin{enumerate}
 \item Zero determinant strategies are not effective invaders or defenders for $N > 2$.
 \item Complex strategies can be effective, and in fact can naturally evolve
     through evolutionary processes to outperform designed strategies.
 \item The strongest resistors specifically evolve or possess a handshake mechanism.
 \item Strong invaders are generally cooperative strategies that do not defect
 first but retaliate to varying degrees of intensity against strategies that defect.
 \item Strategies evolved to maximize their total payoff can be strong invaders
 and achieve mutual cooperation with many other strategies.
\end{enumerate}

@@

All of the training code is archived at \cite{marc_harper_2017_824264}. This
software is (similarly to the Axelrod library) available on
GitHub\footnote{\url{https://github.com/Axelrod-Python/axelrod-dojo}}
with documentation to
train new strategies easily. Training
typically takes less than 100 generations and can be completed within several
hours on commodity hardware.

There are three further strategies trained specifically for this study; Trained
FSM 1, 2, and 3 (TF1 - TF3). These are based on finite state machines of 16, 16,
and 8 states respectively. These are shown in Fig~\ref{fig:tf1},~\ref{fig:tf2}
and~\ref{fig:tf3}, using the notation common in the literature where
\(A_1/A_2\) is the action of the opponent \(A_1\) and the response of the player
\(A_2\) as well as arrows corresponding to changes of state.

@@

As will be demonstrated the results for
\(N=2\) differ from those of larger $N$. Hence these results do not concur with
the literature which suggests that zero determinant strategies should be
effective for larger population sizes, but these analyses consider stationary
behaviour, while this work runs for a fixed number of rounds \cite{stewart2013extortion}.
The stationarity assumption allows for a deterministic payoff matrix
leading to the conclusions about zero determinant strategies in the space
of memory-one strategies that do not generalize to this context.
